On First-order Proofs without syntax

### 2020-12-24 main.tex
Add some more definitions

% ** SPECIAL **

\newcommand*{\Gr}{\mathcal{G}}
\newcommand*{\LG}{\mathcal{L}}

@@

\bfit{Terms} $s, t, u, \ldots$ and \bfit{atoms} (atomic formulas) $\alpha, \beta,
\gamma, \ldots$ are generated inductively from variables $x, y, z, \ldots$: if
$f$ (resp. $p$) is an $n$-ary function (resp. predicate) symbol and $t_1, \ldots,
t_n$ are terms, then $ft_1 \ldots t_n$ (resp. $pt_1 \ldots t_n$) is a term
(resp. atom). The set of atoms is extended with logical constant $\ttt$ and
$\fff$. We also assume that each predicate symbol $p$ has a \bfit{dual}
predicate symbol $\overline{p}$ with $\overline{p} \neq p$ and
$\overline{\overline{p}} = p$. We extend the duality to atoms with
$\overline{pt_1 \ldots t_n} = \overline{p}t_1 \ldots t_n$, $\overline{\ttt} =
\fff$, and $\overline{\fff} = \ttt$. \bfit{Formulas} $A, B, C$ are generated inductively from atoms by binary
connectives $\vee$ and $\wedge$ and quantifiers $\forall x$ and $\exists x$.
The \bfit{negation} $\neg$ is defined in the usual way. A formula is \bfit{rectified} if all bound variables are distinct from one another and from all free variables. Without loss of generality, we assume that all formulas are rectifed throughtout the paper as every formula has a logically equivalent rectified form.

A \bfit{directed graph} $G = (V, E)$ consists of a finite set $V$ of
\bfit{vertices} and a set of $E \subseteq V \times V$ of \bfit{direct edges}. An
\bfit{undirected graph} $G = (V, E)$ consists of a finite set $V$ of vertices
and a set of $E$ of \bfit{undirected edges} (two-element subsets of $V$).
For a given (directed or undirected) graph $G$, we write $V_G$ and $E_G$ for its
vertex and edge sets respectively. We write $vw$ for an edge $(v, w)$ or $\{v,
w\}$. Let $G = (V, E)$ and $G' = (V', E')$ be two undirected graphs. Without
loss of generality, we assume that $V$ and $V'$ are disjoint. The \bfit{union}
$G + G'$ is the graph $(V \cup V', E \cup E')$ and the \bfit{join} $G \times G'$
is the graph $(V \cup V', E \cup E' \cup \{vv' | v \in V, v' \in V'\})$. A
\bfit{cograph} is a $P_4$-free undireced grpah, i.e., for any distinct vertices $v_1,
v_2, v_3, v_4 \in V$, the restriction of the edge sets on them is not equal to
$\{v_1v_2, v_2v_3, v_3v_4\}$. In a graph $G = (V, E)$, the \bfit{neighbourhood}
$N(v)$ of $v \in V$ is defined as the set $\{w | vw \in E\}$, a set $W \subseteq
V$ induces a \bfit{matching} if for all $w \in W$, $N(w) \cap E$ is a singleton, and a module is a
set $M \subseteq V$ such that $N(v) \ M = N(w) \ M$ for all $v, w \in M$. A
\bfit{module} $M$ is strong if for every module $M'$, we have $M' \subseteq M$, $M
\subseteq$ or $M \cap M' = \emptyset$. A \bfit{graph homomorphism} $h: (V, E)
\rightarrow (V', E')$ is a function $h : V \rightarrow V'$ such that $h(v)h(w)
\in E'$ whenever $vw \in E$.

\subsection{First-order Graphs}
A cograph is \bfit{logical} if every vertex is labelled either by an atom or variable,
and it has at least one atom-labelled vertex. We denote by $label_G(v)$, or simply
$label(v)$ the label of a vertex in a logical cograph $G$. An atom-labelled vertex is called
a \bfit{literal} and a variable-labelled vertex is called a \bfit{binder}. A
binder labelled with $x$ is called an \bfit{$x$-binder}. The \bfit{scope} of a binder
$b$ is the smallest strong module containing at least two vertices, including
the binder itself. An \bfit{$x$-literal} is a literal whose atom contains the variable $x$. An $x$-binder \bfit{binds} every $x$-literal in its scope.
In a logical cograph $G$, a binder $b$ is \bfit{existential} (resp. \bfit{universal}) if, for every other
vertex $v$ in its scope, we have $bv \in E_G$ (resp. $bv \notin E_G$). An
$x$-binder is \bfit{legal} if its scope contains no other $x$-binder and at least one
literal.

@@

\subsection{Fonets}
In a fograph, two atoms are \bfit{pre-dual} if their predicate symbols are dual
(e.g. $pxy$ and $\overline{p}yz$) and two literals are \bfit{pre-dual} if their
labels (atoms) are pre-dual. A \bfit{linked fograph} is a coloured fograph such
that every colour, called a \bfit{link}, consists of two pre-dual literals, and
every literal is either $\ttt$-labelled or in a link.

@@

\subsection{Skew bifibrations}

A graph homomorphism $h: (V, E) \rightarrow (V', E')$ is a \bfit{fibration} if
for all $v \in V$ and $wh(v) \in E'$, there exists a unique $\tilde{w}$ such
that $\tilde{w}v \in E$ and $h(\tilde{w}) = w$, and is a \bfit{skew fibration}
if for all $v \in V$ and $wh(v) \in E'$ there exists $\tilde{w}$ such that
$\tilde{w}v \in E$ and $h(\tilde{w})w \notin E'$. A \bfit{fograph homomorphism}
is a graph homomorphism between the underlying graphs of the fographs
considered. A fograph homomorphism $h : G \rightarrow G'$ is
\bfit{label-preserving} if forall $v \in V_G, label(v) = label(h(v))$, and is
\bfit{existential-preserving} if for all existental binder $b$ in $G$, the
vertex $h(b)$ is an existential binder in $G'$.

@@

@@

%\input{bibliography}

>>>>

% ** SYMBOLS **

\newcommand{\ax}{\mathsf{ax}}
\newcommand{\cut}{\mathsf{cut}}
\newcommand{\conj}{\mathsf{\wedge}}
\newcommand{\disj}{\mathsf{\vee}}
\newcommand{\univ}{\mathsf{\forall}}
\newcommand{\exist}{\mathsf{\exists}}
\newcommand{\mix}{\mathsf{mix}}
% ** SPECIAL **

\newcommand{\Gr}{\mathcal{G}}
\newcommand{\LG}{\mathcal{L}}
\newcommand{\FT}{\mathcal{F}}
\newcommand{\US}{\mathcal{U}}

@@

\bfit{Terms} $s, t, u, \ldots$ and \bfit{atoms} (atomic formulas) $\alpha, \beta,
\gamma, \ldots$ are generated inductively from variables $x, y, z, \ldots$: if
$\delta$ (resp. $p$) is an $n$-ary function (resp. predicate) symbol and $t_1, \ldots,
t_n$ are terms, then $\delta t_1 \ldots t_n$ (resp. $pt_1 \ldots t_n$) is a term
(resp. atom). The set of atoms is extended with logical constant $\ttt$ and
$\fff$. We also assume that each predicate symbol $p$ has a \bfit{dual}
predicate symbol $\overline{p}$ with $\overline{p} \neq p$ and
$\overline{\overline{p}} = p$. We extend the duality to atoms with
$\overline{pt_1 \ldots t_n} = \overline{p}t_1 \ldots t_n$, $\overline{\ttt} =
\fff$, and $\overline{\fff} = \ttt$. \bfit{Formulas} $A, B, C$ are generated inductively from atoms by binary
connectives $\vee$ and $\wedge$ and quantifiers $\forall x$ and $\exists x$.
The \bfit{negation} $\neg$ is defined in the usual way. A formula is \bfit{rectified} if all bound variables are distinct from one another and from all free variables. Without loss of generality, we assume that all formulas are rectifed throughtout the paper as every formula has a logically equivalent rectified form.

A \bfit{directed graph} $G = (V, E)$ consists of a finite set $V$ of
\bfit{vertices} and a set of $E \subseteq V \times V$ of \bfit{direct edges}. An
\bfit{undirected graph} $G = (V, E)$ consists of a finite set $V$ of vertices
and a set of $E$ of \bfit{undirected edges} (two-element subsets of $V$).
For a given (directed or undirected) graph $G$, we write $V_G$ and $E_G$ for its
vertex and edge sets respectively. We write $vw$ for an edge $(v, w)$ or $\{v,
w\}$. Let $G = (V, E)$ and $G' = (V', E')$ be two undirected graphs. Without
loss of generality, we assume that $V$ and $V'$ are disjoint. The \bfit{union}
$G + G'$ is the graph $(V \cup V', E \cup E')$ and the \bfit{join} $G \times G'$
is the graph $(V \cup V', E \cup E' \cup \{vv' | v \in V, v' \in V'\})$. A
\bfit{cograph} is a $P_4$-free undireced grpah, i.e., for any distinct vertices $v_1,
v_2, v_3, v_4 \in V$, the restriction of the edge sets on them is not equal to
$\{v_1v_2, v_2v_3, v_3v_4\}$. In a graph $G = (V, E)$, the \bfit{neighbourhood}
$N(v)$ of $v \in V$ is defined as the set $\{w | vw \in E\}$, a set $W \subseteq
V$ induces a \bfit{matching} if for all $w \in W$, $N(w) \cap E$ is a singleton, and a module is a
set $M \subseteq V$ such that $N(v) \ M = N(w) \ M$ for all $v, w \in M$. A
\bfit{module} $M$ is strong if for every module $M'$, we have $M' \subseteq M$, $M
\subseteq$ or $M \cap M' = \emptyset$. A \bfit{graph homomorphism} $f: (V, E)
\rightarrow (V', E')$ is a function $f : V \rightarrow V'$ such that $f(v)f(w)
\in E'$ whenever $vw \in E$.

\subsection{First-order combinatorial proofs}
A cograph is \bfit{logical} if every vertex is labelled either by an atom or variable,
and it has at least one atom-labelled vertex. We denote by $label_G(v)$, or simply
$label(v)$ the label of a vertex in a logical cograph $G$. An atom-labelled vertex is called
a \bfit{literal} and a variable-labelled vertex is called a \bfit{binder}. A
binder labelled with $x$ is called an \bfit{$x$-binder}. The \bfit{scope} of a binder
$b$ is the smallest strong module containing at least two vertices, including
the binder itself. An \bfit{$x$-literal} is a literal whose atom contains the variable $x$. An $x$-binder \bfit{binds} every $x$-literal in its scope.
In a logical cograph $G$, a binder $b$ is \bfit{existential} (resp. \bfit{universal}) if, for every other
vertex $v$ in its scope, we have $bv \in E_G$ (resp. $bv \notin E_G$). An
$x$-binder is \bfit{legal} if its scope contains no other $x$-binder and at least one
literal.

@@

In a fograph, two atoms are \bfit{pre-dual} if their predicate symbols are dual
(e.g. $pxy$ and $\overline{p}yz$) and two literals are \bfit{pre-dual} if their
labels (atoms) are pre-dual. A \bfit{linked fograph} is a coloured fograph such
that every colour, called a \bfit{link}, consists of two pre-dual literals, and
every literal is either $\ttt$-labelled or in a link.

@@

A graph homomorphism $f: (V, E) \rightarrow (V', E')$ is a \bfit{fibration} if
for all $v \in V$ and $wf(v) \in E'$, there exists a unique $\tilde{w}$ such
that $\tilde{w}v \in E$ and $f(\tilde{w}) = w$, and is a \bfit{skew fibration}
if for all $v \in V$ and $wf(v) \in E'$ there exists $\tilde{w}$ such that
$\tilde{w}v \in E$ and $f(\tilde{w})w \notin E'$. A \bfit{fograph homomorphism}
is a graph homomorphism between the underlying graphs of the fographs
considered. A fograph homomorphism $f : G \rightarrow G'$ is
\bfit{label-preserving} if forall $v \in V_G, label(v) = label(f(v))$, and is
\bfit{existential-preserving} if for all existental binder $b$ in $G$, the
vertex $f(b)$ is an existential binder in $G'$.

@@

\begin{definition}
	A \bfit{first-order combinatorial proof} of a fograph $G$ is a skew
bifibration $f: N \rightarrow G$ where $N$ is a fonet. A \bfit{first-order
combinatorial proof} of a formula $A$ is a combinatorial proof of its graph
	$\Gr(A)$.
\end{definition}

\subsection{$\FOMLL$ and Unification Nets}

In $\FOMLL$, terms, atoms, formulas are defined as in first-order logic. For
simplicity, we choose to use $\vee$ and $\wedge$ instead of $\parr$ and
$\otimes$ which are generally used in the presentation of linear logic.
A formula $A$ is identified with its \bfit{formula tree} $\FT(A)$, a directed
tree with leaves labelled by atoms, internal nodes labelled by connectives and
quantifiers, and edges directed from leaves to the root. A \bfit{sequent} $\Gamma$ is simply a disjoint union of formulas. We write
comma for disjoint union.

Sequents are proved using the inference rules of $\FOMLL$:
\begin{center}

\mbox{
\begin{prooftree}
  \infer0[$\ax$]{\vdash A, \neg A}
\end{prooftree}
\hspace{2cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, \neg A}
  \infer2[$\cut$]{\vdash \Gamma, \Delta}
\end{prooftree}}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, B}
  \infer2[$\conj$]{\vdash \Gamma, \Delta, A \wedge B}
\end{prooftree}
\hspace{1.5cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A, B}
  \infer1[$\disj$]{\vdash \Gamma, A \vee B}
\end{prooftree}}

\mbox{
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \infer1[$\univ$]{\vdash \Gamma, \forall x A}
\end{prooftree}
($x \notin fv(\Gamma)$)
\hspace{2cm}

\begin{prooftree}
  \hypo{\vdash \Gamma, A[t/x]}
  \infer1[$\exist$]{\vdash \Gamma, \exists x A}
\end{prooftree}}
\end{center}

We also consider the $\mix$ rule:

\begin{center}
\begin{prooftree}
\hypo{\vdash \Gamma}
\hypo{\vdash \Delta}
\infer2[$\mix$]{\vdash \Gamma, \Delta}
\end{prooftree}
\end{center}

Let $\Gamma$ be a sequent in $\FOMLLm$. A \bfit{link} on $\Gamma$ is a pair of
leaves whose atoms are pre-dual. A \bfit{linking} on $\Gamma$ is a set of 
disjoint links such that each leaf of $\Gamma$ is in exacly one link. Similar
to the set of links in the linked fograph, a linking can be seen as a
unification problem, and a link is said \bfit{unifiable} if the corresponding
unification problem is solvable. \bfit{Dependencies} are defined as previously.

Let $\lambda$ is a unifiable linking on $\Gamma = A_1, \ldots, A_n$.
Let $A = A_1 \vee \ldots \vee A_n$. The \bfit{unification structure}
$\US(\lambda)$
associated to $\lambda$ is the formula tree $\FT(A)$ together with an undirected
edge between leaves $l$ and $l'$ for every link $\{l, l'\}$ in $\lambda$ and a
directed edge from $\exists x$ to $\forall y$ for every dependency $\exists x
\rightarrow \forall y$.

A \bfit{switching graph} of a unification structure $\US(\lambda)$ is any
derivative of $\US(\lambda)$ obtained by keeping only one edge into each $\parr$
and $\forall$ and undirecting remaining edges. A linking is \bfit{correct} if it
is unifiable and all of the switching graphs of its associated unification
structure are acyclic.

\begin{definition}
  A \bfit{unification net} on a sequent $\Gamma$ is a correct linking on $\Gamma$.
\end{definition}

\subsection{Deep Inference}
In this paper we also use some \bfit{deep inference} \cite{BT 2001} rules that are in the
following form:
\begin{center}
\begin{prooftree}
  \hypo{\vdash S\{A\}}
  \infer1[]{\vdash S\{B\}}
\end{prooftree}
\end{center}
where $S\{ \ \}$ stands for a \bfit{context}, which corresponds to a sequent
or a formula with a hole taking the place of an atom, and $S\{A\}$ represents
the sequent or formula obtained by replacing the hole in $S\{ \ \}$ with the formula $A$. Formally,

  \[C ::= \Box \ | \ A \vee C \ | \ C \wedge A \ | \ \exists x C \ | \ \forall x C \].
  \[S ::= C \ | \ A, S \ | \ S, A \] where $A$ is a formula.
The above rule can be thus seen as the rewriting rule $A \rightarrow B$.

We use the notation $\vlderivation{\vlde{}{\mathcal{P}}{B}{\vlhy {A}}}$ for denoting that there is a derivation from premise $\vdash S\{A\}$ to conclusion $\vdash S\{B\}$ in system $P$ for any context $S$.

@@

\input{bibliography}
